---
title: "reproducible-research-rmarkdown"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r}
library(ggplot2)
library(scales) 
```
#Loading and preprocessing the data
```{r}
 
data <- read.csv("activity.csv")
```
#What is mean total number of steps taken per day?
```{r}
library(ggplot2)
total.steps <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
mean(total.steps, na.rm=TRUE)
median(total.steps, na.rm=TRUE)
```
#What is the average daily activity pattern?
```{r}
library(ggplot2)
averages <- aggregate(x=list(steps=data$steps), by=list(interval=data$interval),
                      FUN=mean, na.rm=TRUE)
ggplot(data=averages, aes(x=interval, y=steps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken")
```
On average across all the days in the dataset, the 5-minute interval contains the maximum number of steps?
```{r}
averages[which.max(averages$steps),]
```

#Imputing missing values
 
There are many days/intervals where there are missing values (coded as  NA ). The presence of missing days may introduce bias into some calculations or summaries of the data.
```{r}
missing <- is.na(data$steps)
#  total number of missing values in the dataset
table(missing)
```
All of the missing values are filled in with mean value for that 5-minute interval.

Replace each missing value with the mean value of its 5-minute interval
```{r}
fill.value <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps))
        filled <- c(steps)
    else
        filled <- (averages[averages$interval==interval, "steps"])
    return(filled)
}
filled.data <- data
filled.data$steps <- mapply(fill.value, filled.data$steps, filled.data$interval)
```
Now, using the filled data set, let's make a histogram of the total number of steps taken each day and calculate the mean and median total number of steps.
```{r}
total.steps <- tapply(filled.data$steps, filled.data$date, FUN=sum)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
mean(total.steps)
median(total.steps)
```

Mean and median values are higher after imputing missing data. The reason is that in the original data, there are some days with  steps  values  NA  for any  interval . The total number of steps taken in such days are set to 0s by default. However, after replacing missing  steps  values with the mean  steps  of associated  interval  value, these 0 values are removed from the histogram of total number of steps taken each day.

#Are there differences in activity patterns between weekdays and weekends?

Find the day of the week for each measurement in the dataset.
```{r}
weekday.or.weekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
        return("weekday")
    else if (day %in% c("Saturday", "Sunday"))
        return("weekend")
    else
        stop("invalid date")
}
filled.data$date <- as.Date(filled.data$date)
filled.data$day <- sapply(filled.data$date, FUN=weekday.or.weekend)
```
Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
```{r}
averages <- aggregate(steps ~ interval + day, data=filled.data, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) +
    xlab("5-minute interval") + ylab("Number of steps")
```